Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2023-12-05 manuscript.tex
Update citation format

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano,
Scharler, Scharler, Dunne, \& Ulanowicz, 2005; Thompson et al., 2012).
Trophic networks (or food webs) allow identifying properties and key
species that may be crucial for ecosystem stability, and hence important
for ecosystem management and biodiversity conservation (Thompson et al.,
2012). There is a growing interest in understanding how ecological
networks respond to environmental and anthropogenic stressors along
geographical gradients (Cirtwill, Stouffer, \& Romanuk, 2015; Bauer et
al., 2022). Yet, only a few studies have described variation in food web
structure along latitudinal gradients in marine ecosystems. The few that
have come from the Global North (Wood, Russell, Hanson, Williams, \&
Dunne, 2015; Kortsch et al., 2019; Pecuchet et al., 2022), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

@@

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha, Mianzan, Guerrero, Favero, \& Bava, 2004; Latorre et al.,
2023). The referred region extends from San Jorge Gulf (45ºS) in the
Patagonian shelf to the Weddell Sea (78ºS) in the Southern Ocean, and
covers a well-connected oceanic latitudinal gradient (R. P. Matano,
Palma, \& Piola, 2010; Guihou, Piola, Palma, \& Chidichimo, 2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa \& Varisco, 2007; Pasotti et al.,
2015; Saunders, Hill, Tarling, \& Murphy, 2019; Riccialdelli et al.,
2020), but few studies have considered the complexity of the ecosystem
in terms of a high resolution of species and their prey-predator
interactions (but see Jacob et al. (2011), Marina et al. (2018), Funes,
Saravia, Cordone, Iribarne, \& Galván (2022), López-López, Genner,
Tarling, Saunders, \& O'Gorman (2022), Rodriguez, Marina, Schloss, \&
Saravia (2022), Marina et al. (in rev.)). Neglecting this complexity
might lead to a misunderstanding of the structure and functioning of the
ecosystems, and ultimately reduce the ability to predict ecosystem
responses to perturbations (Montoya, Woodward, Emmerson, \& Solé, 2009).

@@

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Map_final.jpg}
\caption{Map of the study areas along the southwest Atlantic - Antarctic
latitudinal gradient. The areas are marked with numbers from one to six.
Smaller areas (1. San Jorge Gulf, 2. Beagle Channel, 3. Burdwood Bank,
and 5. Potter Cove) are shown on the panels to the left, whereas the
larger areas (4. North and South Scotia Sea, and 6. Weddell Sea) are
marked with a red rectangle on the map. The map was drawn using the
`marmap' R package (Pante, Simon-Bouhet, \& Irisson, 2023). Continental
contour shapefiles were obtained from www.ign.gob.ar.}
\end{figure}

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Glembocki, Williams, Góngora, Gagliardini, \&
Orensanz, 2015). The Gulf's productivity supports large invertebrate and
vertebrate fisheries (Góngora, González-Zevallos, Pettovello, \& Mendía,
2012), as well as marine mammal and seabird populations (Yorio, 2009).
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic
interactions with a connectance of 0.04. The percentage of top predators
is 16\%, 78\% of intermediate nodes and 6\% of basal nodes; 60\% of
predators are omnivorous (Table 1). The most connected nodes are: the
Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea
gregaria, squids (Illex argentinus as dominant species) and Amphipoda.
Notably, these nodes present mid-trophic positions in the food web (3.0,
2.5, 3.6 and 2, respectively) (Funes et al., 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al., 2023). Bruno, Riccialdelli, Acha, \&
Fernández (2023) suggests that locally produced suspended particulate
organic matter (mostly composed by phytoplankton) and organic matter
accumulated in the sediments (macroalgae-originated detritus) are the
primary food sources for the marine ecosystem, as opposed to
allochthonous materials. Beagle Channel food web includes 145 nodes and
1115 trophic interactions with a connectance of 0.05 (Table 1). The food
web is suggested to have a wasp-waist structure, where the following
species play a crucial role in the dynamics of the ecosystem: Fueguian
sprat Sprattus fuegensis, longtail southern cod Patagonotothen ramsayi,
black southern cod P. tessellata, frogmouth Cottoperca trigloides, and
squat lobster Grimothea gregaria (Riccialdelli et al., 2020). Moreover,
the squat lobster has been identified as a species responsible for
linking modules and connecting the entire food web (Rodriguez et al.,
2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Administración de
Parques Nacionales, 2022) (Figure 1). Physical features in the plateau
are fairly stable, with salinity averaging 34 all year round and
temperature ranging between 4 and 8ºC (Acha et al., 2004). The plateau
is surrounded by steep flanks of up to 4000 m depth, protected by the
Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2;
\textasciitilde55º - 56ºS, \textasciitilde58º - 62ºW). Intense upwelling
and mixing occur in relation to the slope, entraining deep nutrient-rich
waters into the photic layer (Ricardo P. Matano, Palma, \& Combes,
2019), and resulting in a fairly homogeneous water column both spatially
and temporally (Ricardo P. Matano et al., 2019). The Burdwood Bank food
web comprises 379 nodes and 1788 interactions, with a connectance of
0.01, and an asymmetric degree distribution (i.e.~most of the species
have a relatively low number of interactions and few species concentrate
most of them). Almost half of the consumers are omnivores (0.48), and
the network displays a small-world pattern (Marina et al., in rev.)
(Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to
the West and by the island complex of the Scotia Arc to the North, East,
and South, with an approximate extension of 1.5 x 106 km2 and a depth
range of 0 - 3000 m (E. j. Murphy et al., 2006) (Figure 1). Its
oceanography is dominated by the Antarctic Circumpolar Current, which is
spatially structured by frontal systems (Whitworth, 1980). The South
Antarctic Circumpolar Current Front subdivides the Scotia Sea into two
biogeographic regions: the Northern Scotia Sea is characterized by
higher and more variable temperatures, and the Southern Scotia Sea by
lower and more stable temperatures and influenced by seasonal sea ice
(Raymond, 2011). The analysis of the Northern and Southern Scotia Sea
food webs shows that the former is relatively more complex than the
latter: with higher species richness (218 vs 192) and interactions
(10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As
expected from a more complex network, the path length is shorter in the
Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea
network displays a greater proportion of omnivores and a lower mean
trophic level (López-López et al., 2022) (Table 1).

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al., 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al., 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (M. L. Quartino \& Boraso de
Zaixso, 2008). The Potter Cove food web includes 110 nodes and 649
interactions, with an asymmetric degree distribution, and a connectance
value of 0.05 (Table 1). It presents a modular structure (groups of
species interact more strongly with each other than with species
belonging to other groups), that's positively associated with stability,
since perturbation can be retained within modules (Rodriguez et al.,
2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al., 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Teixidó, Garrabou, \& Arntz, 2002). The
Weddell Sea food web exhibits a high level of network complexity,
featuring the greatest number of nodes (490) and trophic interactions
(16041) among the analyzed food webs in this review (Table 1). Its
connectance (0.07) and other structural properties are intermediate
compared to the other food webs. Recently, the interaction strengths of
this food web were estimated, revealing the presence of numerous weak
and few strong interactions, which is consistent with findings in other
complex food webs (Nilsson \& McCann, 2016). Notably, this asymmetric
distribution of interaction strength is likely to promote community
persistence.

@@

A stressor is any environmental or anthropogenic variable that causes a
quantifiable change, irrespective of its direction (increase or
decrease), in a biological response (Orr et al., 2020). In the light of
this, the southwest Atlantic - Antarctic marine biota has been and is
currently subjected to a variety of stressors (e.g.~sea warming, glacial
retreat, ice changes, acidification, species invasion, fisheries, and
contamination). Although it is well-known that multiple stressors act in
concert at any given time (e.g.~warming and fishery; acidification and
contaminants) (Gutt et al., 2021), to date stressor assessments have
been performed individually. Moreover, the potential for interactive
effects of two or more stressors (synergy or antagonism) (Côté, Darling,
\& Brown, 2016) is almost unknown for the region (Rowlands et al.,
2021).

@@

San Jorge Gulf experienced several environmental and anthropogenic
stressors (Table S2). Trawl fisheries discard several species and add
new trophic interactions to the food web which resulted in a decrease in
trait variability and the stability of the system (Rincón-Díaz, Bovcon,
Cochia, Góngora, \& Galván, 2021; Funes et al., 2022). Moreover, it
changed the availability of prey to several predators. For example,
Merluccius hubbsi one of the main bycatch species, became prey item to
non-diving seabirds, like the kelp gull Larus dominicanus
(González-Zevallos \& Yorio, 2006) and reef fishes (Funes, Marinao, \&
Galván, 2019). Although juveniles of M. hubbsi are largely the main
bycatch item, 29 other cartilaginous and 69 bony local fish species were
also registered as incidental catch between 2005 and 2014 (Bovcon,
Góngora, Marinao, \& González-Zevallos, 2013; Ruibal Nuñez, 2020). This
level of impact triggered a shift in the functional diversity of the
assemblages homogenizing the trophic function of fishes (Rincón-Díaz et
al., 2021). Other functional changes were a decrease in the maximum
sizes of individual fish, together with a drop in elasmobranchs biomass
and an increase in crustaceans biomass (Funes, 2020). The significant
increase in crustacean biomass was mostly due to the increase in
Pleoticus muelleri and Grimothea gregaria populations (Funes, 2020).
These species rapidly became the most important prey for the most
abundant fishes in the area: M. hubbsi and Mustelus schmitti (Pasti et
al., 2021). However, the above-mentioned effects of trawl fisheries on
the structure and function of the San Jorge Gulf community may have
changed again, since the fisheries ceased activity in 2015 (Annex I,
Resolution CFP No 7/2018), remaining a small trawling artisanal fishery.

Sea warming is another important environmental stressor in the San Jorge
Gulf, because of the southward shifts of northern fish populations to
the Gulf (Galván et al., 2022). The San Jorge Gulf is especially prone
to be affected by climate-driven shifts in species ranges, because it is
located in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech \&
Ehrlich, 2008). In addition to this tropicalization from northern fish
populations, alien species are also documented to affect the demersal
assemblage (fish and macroinvertebrates) in the Gulf (Galván et al.,
2022). Finally, San Jorge Gulf is exposed to urban and industrial
pollution due to an oil monobuoy from which oil manipulation and general
oil transport along the Patagonian coast registered several oil spills
and chronic oil discharges (García-Borboroglu, Boersma, Reyes, \&
Skewgar, 2008). Other marine systems impacted by oil spill showed an
important decrease in marine bird populations (Irons, Kendall, Erickson,
McDonald, \& Lance, 2000), seals and macroalgae (e.g. Paine et al.
(1996)), with examples of cascading effects throughout the food web
(Peterson, 2001).

@@

A main anthropogenic stressor in the Beagle Channel is the introduction
of invasive species (Table S2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially Chinook salmon Oncorhynchus tshawytscha
causes concern. The detection of Chinook salmon in Tierra del Fuego
dates back to April 2006, and its population has been expanding since
then (Nardi, Fernández, Vanella, \& Chalde, 2019). Being a top predator
Chinook salmon can compete with several native species in the Beagle
Channel (Correa \& Gross, 2008), and prey over native species such as
Notothenioids (Eleginops maclovinus, Patagonotothen tessellata, P.
cornucola, P. sima, Paranotothenia magellanica, Harpagifer bispinis),
the Atherinidae Odontesthes smitti and O. nigricans, the Fuegian sprat
Sprattus fuegensis, and larvae of king crabs (Lithodes antarcticus and
Paralomis granulosa) (Fernández, Ciancio, Ceballos, Riva-Rossi, \&
Pascual, 2010). Ciancio, Pascual, Botto, Frere, \& Iribarne (2008)
observed that Chinook salmon in the Southern Patagonian Shelf Ecosystem
Area primarily feed on sprats and display trophic levels comparable to
those of intermediate-sized fish and cephalopod predators species,
showing significant dietary overlap with Magellanic penguins (Spheniscus
magellanicus). Another potential competitor for Chinook salmon in the
Beagle Channel is the Commerson's dolphin (Cephalorhynchus commersonii),
which shares a similar diet (Riccialdelli et al., 2013).

In the Beagle Channel, other anthropogenic stressors include
contaminants like metals, perfluorinated compounds, hydrocarbons, and
microplastics found in animal tissue and sediments. Some studies
reported varying heavy metal levels in sea mussel tissues and sediments,
indicating bioavailability differences (Duarte, Giarratano, Amin, \&
Comoglio, 2011). Other research found similar carbon and carbohydrate
levels in Ushuaia Bay's surface sediments to those in hypertrophic
ecosystems, along with hydrocarbons and heavy metals linked to port and
industrial activities (Gil, Torres, Amin, \& Esteves, 2011).
Commendatore, Nievas, Amin, \& Esteves (2012) identified three distinct
hydrocarbon accumulation zones in the Bay. Llorca et al. (2012) found
high levels of perfluorinated compounds in algae, fish, and guano.
Ferreira, Lo Nostro, Fernández, \& Genovese (2021) showed that black
southern cod in Ushuaia Bay may be exposed to endocrine-disrupting
compounds from urban and industrial pollution. Pérez et al. (2020) and
Ojeda et al. (2021) found microplastics in M. edulis chilensis and
Nacella magellanica, respectively. These studies link pollutants to
Ushuaia (54°48′S 68°18′W) and its port and industrial activities. We
expect the Beagle Channel area closest to the city to be most affected.
Contaminants can spread through the food web, magnifying the risk to
higher-level organisms. Fioramonti, Ribeiro Guevara, Becker, \&
Riccialdelli (2022) studied mercury transfer in three southwest Atlantic
Ocean food webs and found biodilution in the Beagle Channel and Atlantic
coast of Tierra del Fuego webs. However, mercury concentrations rose
with benthivory in these webs, and higher mercury was found in
phytoplankton and Grimothea gregaria. Squat lobsters connect pelagic and
benthic habitats, so any disruption to them could disrupt the food web
(Rodriguez et al., 2022). Dodino et al. (2022) found the highest mercury
levels in Magellanic penguin feathers from offshore colonies in Tierra
del Fuego. Recently, Ushuaia's kelp forests have seen a decrease in
biodiversity and changes in macroalgae composition due to urban
pollution (Kaminsky et al., in prep.).

@@

The Scotia Sea has experienced one of the largest levels of warming
within the polar regions (Whitehouse et al., 2008). Together with the
Southern Annular Mode anomalies this has caused a long-term decrease in
krill abundance; more pronounced in the northern than in the southern
Scotia Sea (E. J. Murphy et al., 2007). Over the past 90 years, the
krill also showed an increase in mean body length (Atkinson et al.,
2019), which may also alter predator-prey interactions and allow
reaching cooler feeding grounds near the seabed, with the potential to
link krill to unexpected predators (Schmidt et al., 2011). Another
consequence of the mentioned environmental stressors, is a change in
krill distribution due to a southward contraction (Atkinson et al.,
2019). It is noteworthy here, that the distributional shift has not
affected all species inhabiting the Northern Scotia Sea, and most
abundant calanoid copepods have maintained their distribution (Tarling,
Ward, \& Thorpe, 2018). Considering the above, food web models of the
Northern Scotia Sea have suggested shifts in the structure and
functioning of the food web, from krill-based to non-krill-based, where
myctophid fishes and squid are playing important roles (Saunders et al.,
2019).

>>>>

The application of a network perspective has emerged as a powerful tool
to tackle the complexity of species interactions, facilitating a better
understanding of the structure and functioning of ecosystems (Belgrano
et al. 2005; Thompson et al. 2012). Trophic networks (or food webs)
allow identifying properties and key species that may be crucial for
ecosystem stability, and hence important for ecosystem management and
biodiversity conservation (Thompson et al. 2012). There is a growing
interest in understanding how ecological networks respond to
environmental and anthropogenic stressors along geographical gradients
(Cirtwill et al. 2015; Bauer et al. 2022). Yet, only a few studies have
described variation in food web structure along latitudinal gradients in
marine ecosystems. The few that have come from the Global North (Wood et
al. 2015; Kortsch et al. 2019; Pecuchet et al. 2022), whereas no
studies, nor meta-analyses, on geographical variation in marine food
webs exist for the Global South (Southern Hemisphere).

@@

Together, the southwest and the Atlantic sector of the Southern Ocean
comprise one of the most biologically productive regions of the world's
oceans (Acha et al. 2004; Latorre et al. 2023). The referred region
extends from San Jorge Gulf (45ºS) in the Patagonian shelf to the
Weddell Sea (78ºS) in the Southern Ocean, and covers a well-connected
oceanic latitudinal gradient (Matano et al. 2010; Guihou et al. 2020).

Throughout this latitudinal gradient, many investigations have been
carried out addressing the trophic ecology of specific species and
prey-predator relationships (Vinuesa and Varisco 2007; Pasotti et al.
2015; Saunders et al. 2019; Riccialdelli et al. 2020), but few studies
have considered the complexity of the ecosystem in terms of a high
resolution of species and their prey-predator interactions (but see
Jacob et al. (2011), Marina et al. (2018), Funes et al. (2022),
López-López et al. (2022), Rodriguez et al. (2022), Marina et al. (in
rev.)). Neglecting this complexity might lead to a misunderstanding of
the structure and functioning of the ecosystems, and ultimately reduce
the ability to predict ecosystem responses to perturbations (Montoya et
al. 2009).

@@

\begin{figure}
\centering
\includegraphics[width=6.27in,height=3.528in]{Figures/Map_final.jpg}
\caption{Map of the study areas along the southwest Atlantic - Antarctic
latitudinal gradient. The areas are marked with numbers from one to six.
Smaller areas (1. San Jorge Gulf, 2. Beagle Channel, 3. Burdwood Bank,
and 5. Potter Cove) are shown on the panels to the left, whereas the
larger areas (4. North and South Scotia Sea, and 6. Weddell Sea) are
marked with a red rectangle on the map. The map was drawn using the
`marmap' R package (Pante et al. 2023). Continental contour shapefiles
were obtained from www.ign.gob.ar.}
\end{figure}

San Jorge Gulf is the northernmost study area considered in this review.
It is a partially enclosed basin spanning approximately 34,000 km2 and
\textasciitilde100 m of maximum depths, located from Cabo dos Bahias
(44°55'S) to Cabo tres Puntas (47°06'S) (Figure 1). The shallower
northern and southern ends of the Gulf (with depths \textasciitilde40 m)
present two prominent frontal systems. These systems are the areas of
highest productivity (Glembocki et al. 2015). The Gulf's productivity
supports large invertebrate and vertebrate fisheries (Góngora et al.
2012), as well as marine mammal and seabird populations (Yorio 2009).
The San Jorge Gulf food web comprises 165 nodes and 1015 trophic
interactions with a connectance of 0.04. The percentage of top predators
is 16\%, 78\% of intermediate nodes and 6\% of basal nodes; 60\% of
predators are omnivorous (Table 1). The most connected nodes are: the
Argentine red shrimp Pleoticus muelleri, the squat lobster Gimothea
gregaria, squids (Illex argentinus as dominant species) and Amphipoda.
Notably, these nodes present mid-trophic positions in the food web (3.0,
2.5, 3.6 and 2, respectively) (Funes et al. 2022).

At a southern latitude, the Beagle Channel (54°52'S, 68°8'W) is an
interoceanic passage located at the southernmost tip of South America,
spanning 240 km in length and 5 km in width (\textasciitilde1200 km2),
and a range depth of 0 - 140 m (Figure 1). It features complex
coastlines, varying bathymetry, a prevailing west-to-east circulation
pattern, and a significant longitudinal gradient of glacial freshwater
discharge (Schloss et al. 2023). Bruno et al. (2023) suggests that
locally produced suspended particulate organic matter (mostly composed
by phytoplankton) and organic matter accumulated in the sediments
(macroalgae-originated detritus) are the primary food sources for the
marine ecosystem, as opposed to allochthonous materials. Beagle Channel
food web includes 145 nodes and 1115 trophic interactions with a
connectance of 0.05 (Table 1). The food web is suggested to have a
wasp-waist structure, where the following species play a crucial role in
the dynamics of the ecosystem: Fueguian sprat Sprattus fuegensis,
longtail southern cod Patagonotothen ramsayi, black southern cod P.
tessellata, frogmouth Cottoperca trigloides, and squat lobster Grimothea
gregaria (Riccialdelli et al. 2020). Moreover, the squat lobster has
been identified as a species responsible for linking modules and
connecting the entire food web (Rodriguez et al. 2022).

The so-called Burdwood Bank ecosystem comprises Marine Protected Areas
Namuncurá - Burdwood Bank I and II, meaning the shallow submarine
plateau named Burdwood Bank with a 200 m isobath boundary, and a deep
slope that reaches 4000 m in depth, respectively (Administración de
Parques Nacionales 2022) (Figure 1). Physical features in the plateau
are fairly stable, with salinity averaging 34 all year round and
temperature ranging between 4 and 8ºC (Acha et al. 2004). The plateau is
surrounded by steep flanks of up to 4000 m depth, protected by the
Marine Protected Area Namuncurá - Burdwood Bank II (32,000 km2;
\textasciitilde55º - 56ºS, \textasciitilde58º - 62ºW). Intense upwelling
and mixing occur in relation to the slope, entraining deep nutrient-rich
waters into the photic layer (Matano et al. 2019), and resulting in a
fairly homogeneous water column both spatially and temporally (Matano et
al. 2019). The Burdwood Bank food web comprises 379 nodes and 1788
interactions, with a connectance of 0.01, and an asymmetric degree
distribution (i.e.~most of the species have a relatively low number of
interactions and few species concentrate most of them). Almost half of
the consumers are omnivores (0.48), and the network displays a
small-world pattern (Marina et al. in rev.) (Table 1).

The Scotia Sea is a deep-sea basin, delimited by the Drake Passage to
the West and by the island complex of the Scotia Arc to the North, East,
and South, with an approximate extension of 1.5 x 106 km2 and a depth
range of 0 - 3000 m (Murphy et al. 2006) (Figure 1). Its oceanography is
dominated by the Antarctic Circumpolar Current, which is spatially
structured by frontal systems (Whitworth 1980). The South Antarctic
Circumpolar Current Front subdivides the Scotia Sea into two
biogeographic regions: the Northern Scotia Sea is characterized by
higher and more variable temperatures, and the Southern Scotia Sea by
lower and more stable temperatures and influenced by seasonal sea ice
(Raymond 2011). The analysis of the Northern and Southern Scotia Sea
food webs shows that the former is relatively more complex than the
latter: with higher species richness (218 vs 192) and interactions
(10008 vs 7241) and a higher connection overall (0.21 vs 0.20). As
expected from a more complex network, the path length is shorter in the
Northern Scotia Sea food web. In the same sense, the Southern Scotia Sea
network displays a greater proportion of omnivores and a lower mean
trophic level (López-López et al. 2022) (Table 1).

In the Antarctic realm, Potter Cove is a \textasciitilde9 km2 fjord with
a depth range of 0 - 200 m located at 25 de Mayo/King George Island
(62°14'S, 58°38'W, South Shetland Islands) on the West Antarctic
Peninsula (Figure 1). The cove, bordered by the Fourcade Glacier, is
divided into three areas: a) the internal cove, a high
glacier-influenced, soft sediment zone with a 50 m maximum depth; b) the
central cove, a mixed substrate area with low meltwater influence and an
80 m maximum depth; and c) the external cove, ice-free for 60 years with
a 185 m maximum depth and rocky substrate (Jerosch et al. 2018). Potter
Cove's high-latitude location results in variable environmental
conditions due to photoperiod length seasonality. Sea ice often covers
this area in winter (Schloss et al. 2012). With low phytoplankton
biomass, macroalgae, and microphytobenthos are likely the primary food
sources for secondary benthic production (Quartino and Boraso de Zaixso
2008). The Potter Cove food web includes 110 nodes and 649 interactions,
with an asymmetric degree distribution, and a connectance value of 0.05
(Table 1). It presents a modular structure (groups of species interact
more strongly with each other than with species belonging to other
groups), that's positively associated with stability, since perturbation
can be retained within modules (Rodriguez et al. 2022).

Located between 74 and 78ºS, the high Antarctic Weddell Sea shelf spans
approximately 450 km from East to West (Jacob et al. 2011) (Figure 1).
The water depth ranges from 200 to 500 m, with shallower regions being
covered by continental ice that forms the coastline along the eastern
and southern parts of the Weddell Sea. Within this shelf area, exists a
complex three-dimensional benthic habitat characterized by substantial
benthic biomasses and an intermediate to high diversity when compared to
benthic boreal communities (Teixidó et al. 2002). The Weddell Sea food
web exhibits a high level of network complexity, featuring the greatest
number of nodes (490) and trophic interactions (16041) among the
analyzed food webs in this review (Table 1). Its connectance (0.07) and
other structural properties are intermediate compared to the other food
webs. Recently, the interaction strengths of this food web were
estimated, revealing the presence of numerous weak and few strong
interactions, which is consistent with findings in other complex food
webs (Nilsson and McCann 2016). Notably, this asymmetric distribution of
interaction strength is likely to promote community persistence.

@@

A stressor is any environmental or anthropogenic variable that causes a
quantifiable change, irrespective of its direction (increase or
decrease), in a biological response (Orr et al. 2020). In the light of
this, the southwest Atlantic - Antarctic marine biota has been and is
currently subjected to a variety of stressors (e.g.~sea warming, glacial
retreat, ice changes, acidification, species invasion, fisheries, and
contamination). Although it is well-known that multiple stressors act in
concert at any given time (e.g.~warming and fishery; acidification and
contaminants) (Gutt et al. 2021), to date stressor assessments have been
performed individually. Moreover, the potential for interactive effects
of two or more stressors (synergy or antagonism) (Côté et al. 2016) is
almost unknown for the region (Rowlands et al. 2021).

@@

San Jorge Gulf experienced several environmental and anthropogenic
stressors (Table S2). Trawl fisheries discard several species and add
new trophic interactions to the food web which resulted in a decrease in
trait variability and the stability of the system (Rincón-Díaz et al.
2021; Funes et al. 2022). Moreover, it changed the availability of prey
to several predators. For example, Merluccius hubbsi one of the main
bycatch species, became prey item to non-diving seabirds, like the kelp
gull Larus dominicanus (González-Zevallos and Yorio 2006) and reef
fishes (Funes et al. 2019). Although juveniles of M. hubbsi are largely
the main bycatch item, 29 other cartilaginous and 69 bony local fish
species were also registered as incidental catch between 2005 and 2014
(Bovcon et al. 2013; Ruibal Nuñez 2020). This level of impact triggered
a shift in the functional diversity of the assemblages homogenizing the
trophic function of fishes (Rincón-Díaz et al. 2021). Other functional
changes were a decrease in the maximum sizes of individual fish,
together with a drop in elasmobranchs biomass and an increase in
crustaceans biomass (Funes 2020). The significant increase in crustacean
biomass was mostly due to the increase in Pleoticus muelleri and
Grimothea gregaria populations (Funes 2020). These species rapidly
became the most important prey for the most abundant fishes in the area:
M. hubbsi and Mustelus schmitti (Pasti et al. 2021). However, the
above-mentioned effects of trawl fisheries on the structure and function
of the San Jorge Gulf community may have changed again, since the
fisheries ceased activity in 2015 (Annex I, Resolution CFP No 7/2018),
remaining a small trawling artisanal fishery.

Sea warming is another important environmental stressor in the San Jorge
Gulf, because of the southward shifts of northern fish populations to
the Gulf (Galván et al. 2022). The San Jorge Gulf is especially prone to
be affected by climate-driven shifts in species ranges, because it is
located in the ecotone between two biogeographic provinces, the
Argentine (30°S - 44°S) and the Magellanic (43°S - 55°S) (Balech and
Ehrlich 2008). In addition to this tropicalization from northern fish
populations, alien species are also documented to affect the demersal
assemblage (fish and macroinvertebrates) in the Gulf (Galván et al.
2022). Finally, San Jorge Gulf is exposed to urban and industrial
pollution due to an oil monobuoy from which oil manipulation and general
oil transport along the Patagonian coast registered several oil spills
and chronic oil discharges (García-Borboroglu et al. 2008). Other marine
systems impacted by oil spill showed an important decrease in marine
bird populations (Irons et al. 2000), seals and macroalgae (e.g. Paine
et al. (1996)), with examples of cascading effects throughout the food
web (Peterson 2001).

@@

A main anthropogenic stressor in the Beagle Channel is the introduction
of invasive species (Table S2). Salmonidae were introduced to Tierra del
Fuego in the 1930s. Especially Chinook salmon Oncorhynchus tshawytscha
causes concern. The detection of Chinook salmon in Tierra del Fuego
dates back to April 2006, and its population has been expanding since
then (Nardi et al. 2019). Being a top predator Chinook salmon can
compete with several native species in the Beagle Channel (Correa and
Gross 2008), and prey over native species such as Notothenioids
(Eleginops maclovinus, Patagonotothen tessellata, P. cornucola, P. sima,
Paranotothenia magellanica, Harpagifer bispinis), the Atherinidae
Odontesthes smitti and O. nigricans, the Fuegian sprat Sprattus
fuegensis, and larvae of king crabs (Lithodes antarcticus and Paralomis
granulosa) (Fernández et al. 2010). Ciancio et al. (2008) observed that
Chinook salmon in the Southern Patagonian Shelf Ecosystem Area primarily
feed on sprats and display trophic levels comparable to those of
intermediate-sized fish and cephalopod predators species, showing
significant dietary overlap with Magellanic penguins (Spheniscus
magellanicus). Another potential competitor for Chinook salmon in the
Beagle Channel is the Commerson's dolphin (Cephalorhynchus commersonii),
which shares a similar diet (Riccialdelli et al. 2013).

In the Beagle Channel, other anthropogenic stressors include
contaminants like metals, perfluorinated compounds, hydrocarbons, and
microplastics found in animal tissue and sediments. Some studies
reported varying heavy metal levels in sea mussel tissues and sediments,
indicating bioavailability differences (Duarte et al. 2011). Other
research found similar carbon and carbohydrate levels in Ushuaia Bay's
surface sediments to those in hypertrophic ecosystems, along with
hydrocarbons and heavy metals linked to port and industrial activities
(Gil et al. 2011). Commendatore et al. (2012) identified three distinct
hydrocarbon accumulation zones in the Bay. Llorca et al. (2012) found
high levels of perfluorinated compounds in algae, fish, and guano.
Ferreira et al. (2021) showed that black southern cod in Ushuaia Bay may
be exposed to endocrine-disrupting compounds from urban and industrial
pollution. Pérez et al. (2020) and Ojeda et al. (2021) found
microplastics in M. edulis chilensis and Nacella magellanica,
respectively. These studies link pollutants to Ushuaia (54°48′S 68°18′W)
and its port and industrial activities. We expect the Beagle Channel
area closest to the city to be most affected. Contaminants can spread
through the food web, magnifying the risk to higher-level organisms.
Fioramonti et al. (2022) studied mercury transfer in three southwest
Atlantic Ocean food webs and found biodilution in the Beagle Channel and
Atlantic coast of Tierra del Fuego webs. However, mercury concentrations
rose with benthivory in these webs, and higher mercury was found in
phytoplankton and Grimothea gregaria. Squat lobsters connect pelagic and
benthic habitats, so any disruption to them could disrupt the food web
(Rodriguez et al. 2022). Dodino et al. (2022) found the highest mercury
levels in Magellanic penguin feathers from offshore colonies in Tierra
del Fuego. Recently, Ushuaia's kelp forests have seen a decrease in
biodiversity and changes in macroalgae composition due to urban
pollution (Kaminsky et al. in prep.).

@@

The Scotia Sea has experienced one of the largest levels of warming
within the polar regions (Whitehouse et al. 2008). Together with the
Southern Annular Mode anomalies this has caused a long-term decrease in
krill abundance; more pronounced in the northern than in the southern
Scotia Sea (Murphy et al. 2007). Over the past 90 years, the krill also
showed an increase in mean body length (Atkinson et al. 2019), which may
also alter predator-prey interactions and allow reaching cooler feeding
grounds near the seabed, with the potential to link krill to unexpected
predators (Schmidt et al. 2011). Another consequence of the mentioned
environmental stressors, is a change in krill distribution due to a
southward contraction (Atkinson et al. 2019). It is noteworthy here,
that the distributional shift has not affected all species inhabiting
the Northern Scotia Sea, and most abundant calanoid copepods have
maintained their distribution (Tarling et al. 2018). Considering the
above, food web models of the Northern Scotia Sea have suggested shifts
in the structure and functioning of the food web, from krill-based to
non-krill-based, where myctophid fishes and squid are playing important
roles (Saunders et al. 2019).
